Viscous Transport in Eroding Porous Media

### 2019-10-12 docs/transport/erosion_transport.tex
edited whole paper and fixed up bibliography according to JFM formatting

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo[inline]{Bryan is here}
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and vorticity at four evenly spaced instances in time.
Initially, the smallest distance between pairs of bodies is $3.29 \times
10^{-4}$, and the smallest distance between the bodies and solid wall is
$4.50 \times 10^{-3}$.  At these distances, a resolution of
approximately $N_\iin = 27,000$ and $N_\out = 18,000$ discretization
points is required to satisfy the $5h$ threshold needed for the
trapezoid rule to achieve machine precision.  By applying the
Barycentric quadrature rule, we stably simulate erosion at the much
coarser resolution of $N_\iin = 256$ and $N_\out = 1024$ discretization
points.

\begin{figure}
\begin{center}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.225\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In addition to the
channels that develop between the bodies and the solid walls, erosion
leads to two main channels through the geometry---one in the top half
and one near the middle.}
\end{figure}

To avoid computing streamlines at every time step during the erosion
process, we compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2}.  The initial porosity is $30.67\%$ and
the initial tortuosity is $T = 1.24$.  In
figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
This model outperformed the other three models in
equation~\eqref{eqn:tortuosityModels}, and its root-mean-squared error
is $1.13 \times 10^{-2}$.  As grains erode, there is an increase in the
number of streamlines that take a nearly direct path through the
geometry, and this decreases the tortuosity. However, the channelization
effect of erosion results in an increase in the tortuosity since the
length of many of the streamlines increases when they deflect from a
high porosity region (low pressure) to a low porosity region (high
pressure). For this example, we see that the net effect is a decrease in
the tortuosity for all time.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
\label{sec:Eroding100}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_50}
\caption{100 bodies, $\phi = 55.54\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_100}
\caption{99 bodies, $\phi = 62.98\%$}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_150}
\caption{94 bodies, $\phi = 72.22\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_200}
\caption{82 bodies, $\phi = 83.67\%$}
\end{subfigure}
\caption{\label{fig:Eroding100vort} The erosion of 100 nearly touching
grains in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity.  Because of the large
number of bodies, erosion creates many channels connecting the inlet to
the outlet.}
\end{figure}

As a final example, we consider 100 eroding bodies, each with
$N_\iin=256$ discretization points, with an initial porosity near 50\%.
Snapshots of the configurations and vorticity are in
figure~\ref{fig:Eroding100vort}. We compute the tortuosity using both
the Lagrangian~\eqref{eqn:tortuosity1} and Eulerian
methods~\eqref{eqn:tortuosity2}. Therefore, we must compute the velocity
at $N_p = 1000$ points along the inlet $x=-1$. These normalized
velocities are plotted in figure~\ref{fig:Eroding100tort}(a) for the
eroded geometry at porosity $\phi = 62.98\%$
(figure~\ref{fig:Eroding100tort}(c)).  The initial velocity of the
tracers is qualitatively similar to the 20 body example
(figure~\ref{fig:Eroding20tort}(a)), except with additional maximum and
minimums because of the additional bodies.  Next, in
figure~\ref{fig:Eroding100tort}(b), we plot the local tortuosity by
finding the length of the streamlines as they pass from $x=-1$ to $x=1$.
Compared to figure~\ref{fig:Eroding20tort}(b), the local tortuosity is
much more discontinuous.  These discontinuities can be explained by
examining the trajectories of tracers in
figure~\ref{fig:Eroding100tort}(c).  Here, there are many instances of
nearby streamlines that are deflected apart from one another as they
tend to a stagnation point in the flow, and this results in trajectories
with significantly different lengths.  At this porosity, one of the
tracers travels 25.5\% farther than it would have if the bodies had been
absent, and the average tracer travelled 12\% farther indicating a
tortuosity of $1.12$.

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 62.98\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
the 20 body example in figure~\ref{fig:Eroding20tort}.  (b) The local
tortuosity $\tau(y)$ on the cross section $x = -1$. The spatially
averaged tortuosity through the channel is $1.12$. (c) The trajectories
of 200 tracers initialized at $x = -1$.  The discontinuities in the
local tortuosity correspond to nearby streamlines whose lengths are
significantly different because they pass opposite sides of an eroding
body. Compared to figure~\ref{fig:Eroding20tort}, this example has more
small bodies, and this results in many more discontinuities.}
\end{figure}

In figure~\ref{fig:Eroding100Transport}(a), we plot the tortuosity as a
function of the porosity.  The initial geometry has a porosity of $\phi
= 50.09\%$ and the tortuosity is $T = 1.20$.  The tortuosity is computed
with both the length of the streamlines~\eqref{eqn:tortuosity1} (red
stars) and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding100tort}(c).
Again, the two tortuosity formulas give similar results.  For this
geometry, the tortuosity decreases monotonically at almost all
porosities.  However, the tortuosity undergoes a sudden increase near
the end of the simulation, and we have observed this behavior in other
examples.  The increase is caused by a single small body near the middle
of the channel being completely eroded.  While this results in
straighter streamlines, therefore reducing $\lambda$, the horizontal
flow, $u_1(y)$ increases since there is no longer a no-slip boundary,
and this increases the tortuosity.  We also computed the lines of best
fit using the porosity-tortuosity models~\eqref{eqn:tortuosityModels}.
The black dashed line in figure~\ref{fig:Eroding100Transport}(a) is the
line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2459},
\end{align}
with a root-mean-square error of $5.50 \times 10^{-3}$.  We note a
slightly better root-mean-square error of $5.20 \times 10^{-3}$ is
possible with the model
\begin{align}
  \widehat{T}(\phi) = 1 - 0.2631 \ln(\phi).
\end{align}

In figure~\ref{fig:Eroding100Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$ at six different
porosities. Again, we initially observe ballistic motion (black dashed
line), and then super-dispersion.  Similar to the example in
figure~\ref{fig:ErodingLow20Transport}(b), the anomalous dispersion rate
is not growing with the porosity.  Therefore, it appears that the
dispersion rate in an eroding geometry depends not only the porosity,
but also the location and shape of the bodies.  Finally, at the highest
porosity, anomalous dispersion is only observed briefly in the time
interval $(0.5,1)$, and then transitions back to a ballistic regime.
Since the bodies are so small at this high porosity, after reinsertion,
the streamline is not significantly deflected by any of the bodies, and
this results in a ballistic regime.

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/100b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100Transport} (a) The tortuosity of an
eroding geometry initialized with 100 grains.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The dash
line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2459$,
and the root mean square deviation is $5.50 \times 10^{-3}$.  We note
that the model $\widehat{T}(\phi) = 1 - p \ln (\phi)$, with $p=0.2631$,
has a comparable root-mean-square error of $5.20 \times 10^{-3}$.  (b)
The temporal evolution of $\sigma_\lambda$ as a function of time at six
different porosities.  The dashed line, which has slope one, corresponds
to ballistic dispersion $\sigma_\lambda \sim t$ that occurs if no bodies
are present.  At early times, the particles undergo a ballistic motion,
but once they have traversed a few grains, the spreading becomes
super-dispersive with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in
(1/2,1)$.  The dashed-dotted lines are lines of best fit with slope
$1.11$ ($\phi=95.12\%$), $0.68$ ($\phi=85.07\%$), $0.73$
($\phi=75.14\%$), $0.59$ ($\phi=65.03\%$), $0.61$ ($\phi=55.02\%$), and
$0.70$ ($\phi=50.09\%$).}
\end{figure}

Finally, we investigate the effects of erosion on pore sizes.  The
distribution of the pore sizes can be directly related to the
distribution of the velocity, and thus effects the
tortuosity~\citep{den-ica-hid2018} and anomalous
dispersion~\citep{dea-qua-bir-jua2018}. In addition, the pore sizes are
required in network models~\citep{bry-mel-cad1993, bry-kin-mel1993}.  As
described in section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the pore size by
finding the closest distance between all neighboring bodies.  Instead of
computing a Delanuay configuration at each time step, which would result
in new definitions for the pores at each time step, we only compute a
new Delaunay triangulation when a grain completely erodes.  Once all
pore sizes are computed, we analyze their distribution as a function of
the porosity.

In figure~\ref{fig:Eroding100gap_hist}, we plot histograms of the pore
pore sizes at six porosities during the erosion process. These types of
distributions have been described with Weibull
distributions~\citep{ioa-cha1993}, and we superimpose the Weibull
distribution with the same first two moments as the data.  The
parameters of the distribution, $(k,\lambda)$, are included in the
caption of figure~\ref{fig:Eroding100gap_hist}.  In
figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the pore sizes as a function of the porosity.  Interestingly, for
porosities less than $\phi = 85\%$, the mean pore size grows linearly
and the variance remains nearly flat. Since a channelized geometry has
large variance, this indicates that channelization is less prevalent at
low porosities.

\begin{figure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 pores}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 pores}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the pore
sizes of 100 eroding bodies at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840, 0.3194)$.}
\end{figure}

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/gap_variance}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_mean_var} The effect of erosion on the
mean (a) and the variance (b) of the pore sizes. The geometry initially
contains 100 eroding bodies.  The distributions of the pore sizes in
figure~\ref{fig:Eroding100gap_hist} are indicated by the red stars.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a continuation of our previous work~\citep{qua-moo2018}, we have
simulated dense suspensions and characterized transport in viscous
erosion of porous media. A BIE formulation is used to solve the Stokes
equations and to compute the shear stress. However, to simulate
nearly-touching grains and to characterize transport, it is necessary to
integrate near-singular functions. This is achieved by extending a
Barycentric quadrature rule~\citep{bar-wu-vee2015} to accurately compute
the erosion rate and vorticity. By using this quadrature rule, we are
able to perform stable simulations of erosion with $N = O(100)$
discretization points, while the trapezoid rule would require $O(10^5)$
discretization points.

The transport is characterized in terms of tortuosity and anomalous
dispersion. The local tortuosity agrees qualitatively with other
works~\citep{mat-kha-koz2008}. However, the tortuosity of the eroded
geometries cannot be completely described in terms of the porosity. In
particular, we observe that for certain porous media, the tortuosity can
increase when the porosity increases under erosion. We also observe
super-dispersive spreading, and the rate of dispersion significantly
depends not only on the porosity, but also the number of eroding bodies
and their distribution.

@@

As a long term goal, we hope to include the inertial effects and other
transport models. Including inertia requires an integral equation
formulation of the Navier-Stokes equations, which is an active area of
research with a few promising directions recently
proposed~\citep{gray2019boundary, kli-ask-kro2019}.  Regarding other
transport models, this would involve a diffusive term to consider the
transport of heat or a contaminant. Forming high-fidelity simulations of
such an advection-diffusion equation can be accomplished by using time
splitting methods and recent work on heat solvers in complex
geometries~\citep{fry-kro-tor2019}.

%\todo[inline]{dump of text}
%As in our previous work~\citep{qua-moo2018}, we consider two-dimensional
%immobile eroding grains. However, this work does not limit the proximity
%of  grains. Compared to our previous work, we are able to simulate much
%denser suspensions with a comparable number of discretization points and
%without upsampling.  While we do investigate transport in this work, we
%only consider passive particles that are advected by the flow.  To model
%other physical processes, such as contaminant transport, we plan to
%simulate a finite Peclet number advection-diffusion equation in eroded
%geometries.
%\todo[inline]{dump of text}

@@

\bibliographystyle{jfm} 
%\biboptions{sort&compress}
\bibliography{refs}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Low Porosity}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider a second example with 20 eroding bodies, but with a smaller
initial porosity.  In figure~\ref{fig:ErodingLow20vort}, we plot the
eroding geometry and vorticity at four evenly spaced instances in time.
Initially, the smallest distance between pairs of bodies is $3.29 \times
10^{-4}$, and the smallest distance between the bodies and solid wall is
$4.50 \times 10^{-3}$.  At these distances, a resolution of
approximately $N_\iin = 27,000$ and $N_\out = 18,000$ discretization
points is required to satisfy the $5h$ threshold needed for the
trapezoid rule to achieve machine precision.

\begin{figure}
\begin{center}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort1}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort150}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort300}
  \includegraphics[height=0.227\textwidth]{./figs/20b2_vort450}
\end{center}
\caption{\label{fig:ErodingLow20vort} The erosion of 20 nearly touching
bodies in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity. In addition to the
channels that develop between the bodies and the solid walls, erosion
leads to two main channels through the geometry---one in the top half
and one near the middle.}
\end{figure}

We compute the tortuosity using the Eulerian
method~\eqref{eqn:tortuosity2} at each time step. The initial porosity
is $\phi = 30.67\%$ and the initial tortuosity is $T = 1.24$.  In
figure~\ref{fig:ErodingLow20Transport}(a), we plot the tortuosity with
respect to the porosity (blue) and the line of best fit (black) using
the power law 
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.1669}.
\end{align}
This model outperforms the other three models in
equation~\eqref{eqn:tortuosityModels}, and its root-mean-squared error
is $1.13 \times 10^{-2}$.  As grains erode, there is an increase in the
number of streamlines that take a nearly direct path through the
geometry, and this decreases the tortuosity. However, the channelization
effect of erosion results in an increase in the tortuosity since the
length of many of the streamlines increases when they deflect from a
high porosity region (low pressure) to a low porosity region (high
pressure). For this example, we see that the net effect is a decrease in
the tortuosity for all time.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{100 eroding bodies}
\label{sec:Eroding100}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a final example, we consider 100 eroding bodies with an initial
porosity near 50\%.  Snapshots of the configurations and vorticity are
in figure~\ref{fig:Eroding100vort}. We compute the tortuosity using both
the Lagrangian~\eqref{eqn:tortuosity1} and Eulerian
methods~\eqref{eqn:tortuosity2}. Therefore, we compute and plot the
normalized velocity
at $N_p = 1000$ points along the inlet $x=-1$ in figure~\ref{fig:Eroding100tort}(a) for the
eroded geometry at porosity $\phi = 62.98\%$
(figure~\ref{fig:Eroding100tort}(c)).  The initial velocity of the
tracers is qualitatively similar to the 20 body example
(figure~\ref{fig:Eroding20tort}(a)), except with additional oscillations
because of the additional grains.  In
figure~\ref{fig:Eroding100tort}(b), we plot the local tortuosity by
finding the length of the streamlines as they pass from $x=-1$ to $x=1$.
Compared to figure~\ref{fig:Eroding20tort}(b), the local tortuosity is
much more discontinuous.  These discontinuities can be explained by
examining the trajectories of tracers in
figure~\ref{fig:Eroding100tort}(c).  Here, there are many instances of
nearby streamlines that are deflected apart from one another as they
tend to a stagnation point in the flow, and this results in trajectories
with significantly different lengths.  At this porosity, one of the
tracers travels 25.5\% farther than it would have if the bodies had been
absent, and the average tracer travelled 12\% farther resulting in a
tortuosity of $T = 1.12$.

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_50}
\caption{100 bodies, $\phi = 55.54\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_100}
\caption{99 bodies, $\phi = 62.98\%$}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_150}
\caption{94 bodies, $\phi = 72.22\%$}
\end{subfigure}%
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height =0.8\linewidth]{./figs/100b_200}
\caption{82 bodies, $\phi = 83.67\%$}
\end{subfigure}
\caption{\label{fig:Eroding100vort} The erosion of 100 nearly touching
grains in a Hagen-Poiseuille flow. The four snapshots are evenly spaced
in time, and the color is the fluid vorticity.  Because of the large
number of bodies, erosion creates many channels connecting the inlet to
the outlet.}
\end{figure}

In figure~\ref{fig:Eroding100Transport}(a), we plot the tortuosity as a
function of the porosity.  The initial geometry has a porosity of $\phi
= 50.09\%$ and the tortuosity is $T = 1.20$.  The tortuosity is computed
with both the length of the streamlines~\eqref{eqn:tortuosity1} (red
stars) and using the spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2} (blue marks).  The red square corresponds
to the porosity of the geometry in figure~\ref{fig:Eroding100tort}(c).
Again, the two tortuosity formulas give similar results.  For this
geometry, the tortuosity decreases monotonically at almost all
porosities.  However, the tortuosity undergoes a sudden increase near
the end of the simulation, and we have observed this behavior in other
examples.  The increase is caused by a single small body near the middle
of the channel being completely eroded.  While this results in
straighter streamlines, therefore reducing $\lambda$, the horizontal
flow, $u_1(y)$, increases since there is no longer a no-slip boundary,
and this increases the tortuosity.  We also compute the lines of best
fit using the porosity-tortuosity models~\eqref{eqn:tortuosityModels}.
The black dashed line in figure~\ref{fig:Eroding100Transport}(a) is the
line of best fit
\begin{align}
  \widehat{T}(\phi) = \phi^{-0.2459},
\end{align}
with a root-mean-square error of $5.50 \times 10^{-3}$.  We note a
slightly better root-mean-square error of $5.20 \times 10^{-3}$ is
possible with the model
\begin{align}
  \widehat{T}(\phi) = 1 - 0.2631 \ln(\phi).
\end{align}

\begin{figure}
\begin{subfigure}[b]{0.45\textwidth}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/velocity_loc100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{\textwidth}
\includegraphics*[width =\linewidth]{./figs/tort_local100}
\caption{}
\end{subfigure}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics[width = \textwidth]{./figs/100b_t100tracer}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100tort} The tortuosity of a porous geometry
that initially had 100 grains and has eroded to a porosity of 62.98\%.
(a) The $x$-component of the velocity at the inlet, $u_1(-1, y)$,
normalized by its maximum velocity $u_{max}=3.90 \times 10^{-4}$.  Note
that this maximum velocity is about a order of magnitude smaller than
the 20 body example in figure~\ref{fig:Eroding20tort}.  (b) The local
tortuosity $\tau(y)$ on the cross section $x = -1$. The spatially
averaged tortuosity through the channel is $T = 1.12$. (c) The
trajectories of 200 tracers initialized at $x = -1$.  The
discontinuities in the local tortuosity correspond to nearby streamlines
whose lengths are significantly different because they pass opposite
sides of an eroding body. Compared to figure~\ref{fig:Eroding20tort},
this example has more small bodies, and this results in many more
discontinuities.}
\end{figure}

In figure~\ref{fig:Eroding100Transport}(b), we plot the temporal
evolution of the particle spreading $\sigma_\lambda$ at six different
porosities. Again, we initially observe ballistic motion (black dashed
line), and then super-dispersion.  Similar to the example in
figure~\ref{fig:ErodingLow20Transport}(b), the asymptotic anomalous
dispersion rate is not growing with the porosity.  Therefore, it appears
that the dispersion rate in an eroding geometry depends not only the
porosity, but also the location and shape of the bodies.  Finally, at
the highest porosity, anomalous dispersion is only observed briefly in
the time interval $(0.5,1)$, and then transitions back to a ballistic
regime.  Since the bodies are so small at this high porosity, after
reinsertion, the streamline is not significantly deflected by any of the
bodies, and this results in a ballistic regime.

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/tort_eulerian100}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/100b_second_moment_long_ref}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100Transport} (a) The tortuosity of an
eroding geometry initialized with 100 grains.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The dash
line is the fitting line $\widehat{T}(\phi)=\phi^{-p}$ with $p=0.2459$,
and the root-mean-square error is $5.50 \times 10^{-3}$.  We note that
the model $\widehat{T}(\phi) = 1 - p \ln (\phi)$, with $p=0.2631$, has a
comparable root-mean-square error of $5.20 \times 10^{-3}$.  (b) The
temporal evolution of $\sigma_\lambda$ as a function of time at six
different porosities.  The dashed line, which has slope one, corresponds
to ballistic dispersion $\sigma_\lambda \sim t$ that occurs if no bodies
are present.  At early times, the particles undergo a ballistic motion,
but once they have traversed a few grains, the spreading becomes
super-dispersive with $\sigma_\lambda \sim t^{\alpha}$, $\alpha \in
(1/2,1)$.  The dashed-dotted lines are lines of best fit with slope
$1.11$ ($\phi=95.12\%$), $0.68$ ($\phi=85.07\%$), $0.73$
($\phi=75.14\%$), $0.59$ ($\phi=65.03\%$), $0.61$ ($\phi=55.02\%$), and
$0.70$ ($\phi=50.09\%$).}
\end{figure}

Finally, we investigate the effect of erosion on pore sizes.  The
distribution of the pore sizes is directly related to the
distribution of the velocity, and thus effects the
tortuosity~\citep{den-ica-hid2018} and anomalous
dispersion~\citep{dea-qua-bir-jua2018}. In addition, the pore sizes are
used in network models~\citep{bry-mel-cad1993, bry-kin-mel1993}.  As
described in section~\ref{sec:throats}, we use a Delaunay triangulation
to define neighboring eroding bodies, and we compute the pore size by
finding the closest distance between all neighboring bodies.  Instead of
computing a Delanuay configuration at each time step, which would result
in new definitions for the pores at each time step, we only compute a
new Delaunay triangulation when a grain completely erodes.  Once all
pore sizes are computed, we analyze their distribution as a function of
the porosity.

In figure~\ref{fig:Eroding100gap_hist}, we plot histograms of the pore
pore sizes at six porosities throughout the erosion process. We
superimpose the Weibull distribution~\citep{ioa-cha1993} with the same
first two moments as the data.  The parameters of the distribution,
$(k,\lambda)$, are included in the caption of
figure~\ref{fig:Eroding100gap_hist}.  In
figure~\ref{fig:Eroding100gap_mean_var}, we plot the mean and variance
of the pore sizes as a function of the porosity.  Interestingly, for
porosities less than $\phi = 85\%$, the mean pore size grows linearly
and the variance remains nearly flat. Since a channelized geometry has
large variance, this indicates that channelization is less prevalent at
low porosities.

\begin{figure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 pores}
\end{subfigure}
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.33\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 pores}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the pore
sizes of 100 eroding bodies at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840,0.3194)$.}
\end{figure}

\begin{figure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height = 0.8\linewidth]{./figs/gap_mean}
\caption{}
\end{subfigure}
\begin{subfigure}[b]{0.5\textwidth}
\includegraphics*[height=0.8\linewidth]{./figs/gap_variance}
\caption{}
\end{subfigure}
\caption{\label{fig:Eroding100gap_mean_var} The effect of erosion on (a)
the mean and (b) the variance of the pore sizes. The geometry initially
contains 100 eroding bodies. The distributions of the pore sizes in
figure~\ref{fig:Eroding100gap_hist} are indicated by the red stars.}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
As a continuation of our previous work~\citep{qua-moo2018}, we have
simulated dense suspensions and characterized transport in viscous
eroding porous media. This is accomplished by using high-order time
stepping methods and a new quadrature methods to solve a BIE formulation
of the Stokes equations.  By using these numerical methods, we are able
to perform stable simulations of erosion with $N = O(100)$
discretization points, while the trapezoid rule would require $O(10^5)$
discretization points.

The transport is characterized in terms of tortuosity and anomalous
dispersion. While the local tortuosity agrees qualitatively with other
works~\citep{mat-kha-koz2008}, the tortuosity of eroded geometries
cannot be completely described in terms of the porosity. In particular,
we observe that for certain porous media, the tortuosity can increase as
erosion increases the porosity of the porous media. We also observe
super-dispersive spreading, and the rate of dispersion significantly
depends not only on the porosity, but also the number of eroding bodies
and their distribution.

@@

As a long term goal, we plan to include the inertial effects and other
transport models. Including inertia requires an integral equation
formulation of the Navier-Stokes equations, which is an active area of
research with promising directions recently
proposed~\citep{gray2019boundary, kli-ask-kro2019}.  Regarding other
transport models, this would involve a diffusive term to consider the
transport of heat or a contaminant. Forming high-fidelity simulations of
such an advection-diffusion equation can be accomplished by using time
splitting methods and recent work on heat solvers in complex
geometries~\citep{fry-kro-tor2019}.

@@

\bibliographystyle{jfm} 
\bibliography{refs}
